NILPOTENT INVARIANTS \\ FOR GENERIC DISCRETE SERIES OF REAL GROUPS

### 2025-03-06 whittaker_JIMJ.tex
Small changes after Jeff's round

The purpose
of this paper is to give a self-contained account of these matters. More precisely, we will do two things:
\begin{enumerate}
\item Give an elementary proof that  passage from~$\pi$ to $\mathfrak{w}(\pi)$, $\WF(\pi)$ and $\AV(\pi)$ defines bijections between generic discrete series in an $L$-packet for $G(\R)$, Whittaker data for $G(\R)$, and the appropriate sets of nilpotent orbits;
\item Given one of the invariants $\Wh(\pi)$, $\WF(\pi)$ or $\AV(\pi)$, explain in the simplest possible terms how to explicitly reconstruct the other two.
\end{enumerate}
Many results in this note were known. We gave new and simplified proofs for as many of them as we could; for instance, Section~\ref{sec:Sekiguchi} contains a simple proof that $\WF(\pi)$ and $\AV(\pi)$ are exchanged by the Kostant--Sekiguchi correspondence. Otherwise we give convenient references.

@@

\subsubsection{Involutions}~
 By definition there exists an anti-holomorphic involution~$\sigma$ of~$G$ such that $G(\R) = G^\sigma$. Fix such an involution~$\sigma$, and choose a Cartan involution $\theta$ for $G$. This means:
$\theta$ is an algebraic involution, commuting with~$\sigma$,
and $G(\R)^\theta$ is a maximal compact subgroup of~$G(\R)$. Set $K=G^\theta$ and $K(\R)=G(\R)^\theta=K^\sigma$.
Set $\g=\Lie(G)$, and let $\s$ be the $(-1)$-eigenspace of $\theta$ acting on~$\g$, where we use the notation $\sigma, \theta$ also for the automorphisms of~$\g$ induced by the above involutions. Set $\g(\R)=\Lie(G(\R))$.

@@

Recall $G(\R)$ has relative discrete series representations if and only if it has a relatively compact Cartan subgroup, i.e. there exists a $\theta$-stable Cartan subgroup~$T$ of~$G$  such that~$T(\R) \cap G_d$ is  compact, where $G_d$ is the derived group of~$G$. Throughout the paper, we assume $G(\R)$ does have discrete series representation, we fix  a Cartan subgroup $T \subset G$ satisfying the conditions above, and we let~$\mathfrak{t}$ denote the Lie algebra of~$T$, which is a Cartan subalgebra of~$\g$.

@@

@@

\section{Appendix: the case of $SL(2,\R)$}

>>>>

The purpose
of this paper is to give a self-contained account of these matters. More precisely, we will do two things:
\begin{enumerate}
\item Give an elementary proof that  passage from~$\pi$ to $\mathfrak{w}(\pi)$, $\WF(\pi)$ and $\AV(\pi)$ defines bijections between generic discrete series in an $L$-packet for $G(\R)$, Whittaker data for $G(\R)$, and the appropriate sets of nilpotent orbits;
\item Given one of the invariants $\Wh(\pi)$, $\WF(\pi)$ or $\AV(\pi)$, explain in the simplest possible terms how to explicitly reconstruct the other two.
\end{enumerate}
Many results in this note are known. We gave new and simplified proofs for as many of them as we could; for instance, Section~\ref{sec:Sekiguchi} contains a simple proof that $\WF(\pi)$ and $\AV(\pi)$ are exchanged by the Kostant--Sekiguchi correspondence. Otherwise we give convenient references.

@@

\subsubsection{Involutions}~
 In our setting we may view $G(\R)$ as the fixed-point group $G^\sigma$, where $\sigma$ is an anti-holomorphic involution~$\sigma$ of~$G$ such that $G(\R) = G^\sigma$. Fix such an involution~$\sigma$, and choose a compatible Cartan involution~$\theta$ for $G$. This means:
$\theta$ is an algebraic involution, commuting with~$\sigma$,
and $G(\R)^\theta$ is a maximal compact subgroup of~$G(\R)$. Set $K=G^\theta$ and $K(\R)=G(\R)^\theta=K^\sigma$.
Set $\g=\Lie(G)$, and let $\s$ be the $(-1)$-eigenspace of~$\theta$ acting on~$\g$, where we use the notation $\sigma, \theta$ also for the automorphisms of~$\g$ induced by the above involutions. Set $\g(\R)=\Lie(G(\R))$.

@@

Recall $G(\R)$ has relative discrete series representations if and only if it has a relatively compact Cartan subgroup, i.e. there exists a $\theta$-stable Cartan subgroup~$T$ of~$G$  such that~$T(\R) \cap G_d$ is  compact, where $G_d$ is the derived group of~$G$. Throughout the paper, we assume $G(\R)$ does have relative discrete series representation, we fix  a Cartan subgroup $T \subset G$ satisfying the conditions above, and we let~$\mathfrak{t}$ denote the Lie algebra of~$T$, which is a Cartan subalgebra of~$\g$.

@@

We encourage the reader to consult the example of $\mathrm{SL}(2,\R)$ in the Appendix (\S\,\ref{sec:appendix}) when reading this section.

@@

\section{Appendix: the case of $SL(2,\R)$}\label{sec:appendix}
